Fast Regression of the Tritium Breeding Ratio in Tokamak Fusion Reactors

### 2020-05-25 paper/results.tex
shorten the first two results sections, transition to two-column layout

\todo{This entire section needs to be shortened.}

Having outlined a variety of models and metrics tracked for the
purposes of their objective comparison, we proceed to present and discuss our
results in the next sections.

\subsection{Results of Decoupled Sampling}
\label{sec:modelres}

We begin by comparing the performance of a diverse set of surrogate families on previously generated samples of the expensive MC TBR model.
Through the four experimental cases described
in~\sref{sec:experiment-methodology}, we aim to study properties of the
considered models in terms of regression performance, training and prediction
time.

@@

The first two experiments perform Bayesian optimisation to maximise~$R^2$ in
a cross-validation setting as a function of model hyperparameters. While in the
first experiment we limit training and test sets to the scope of four selected
slices of the feature space, in the second experiment we lift this restriction
to examine surrogate capability to model a more complex domain.

The results displayed in~\fref{fig:exp1-time-vs-reg} (and listed
in~\tref{tbl:exp1-detailed-results} in the Appendix) indicate that in the first
experiment, GBTs clearly appear to be the most accurate as
well as the fastest surrogate family in terms of mean prediction time. Following
that, we note that ERTs, SVMs and ANNs also achieved satisfactory results with respect to both examined metrics.
While the remainder of tested surrogate families does not exhibit problems in
complexity, its regression performance falls below average.

\begin{figure}[h]
	\centering
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{exp1_slice0}
		%\caption{Run 2, batches 0-2}
	\end{minipage}\hfill%
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{exp1_slice1}
		%\caption{Run 2, batches 100-102}
	\end{minipage}\hfill%
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{exp1_slice2}
		%\caption{Run 2, batches 200-202}
	\end{minipage}
	\caption{20~best-performing surrogates per each considered family, plotted in
		terms of complexity (as~$\overline{t}_{\text{pred.}}$) and regression
		performance (as~$R^2$) on selected slices of run~2, evaluated in
	experiment~1. Here, batches refer to subsets of training and test datasets that
	may be matched to slices using~\tref{tbl:slices}.}
	\label{fig:exp1-time-vs-reg}
\end{figure}

The results of the second experiment, shown in~\fref{fig:exp2-time-vs-reg} (and listed
in~\tref{tbl:exp2-detailed-results} in the Appendix),
seem to confirm our expectations. Compared to the previous case, we observe
that many surrogate families consistently achieved worse regression
performance and prediction times. The least
affected models appear to be GBTs, ANNs and ERTs, which are known to be capable of capturing relationships
involving mixed feature types that were deliberately withheld in the first
experiment. With only negligible differences, the first two of these families
appear to be tied for the best performance as well as the shortest prediction
time. We observe that ERTs and RBFs also
demonstrated satisfactory results, relatively outperforming the remaining surrogates in
terms of regression performance, and in some cases also in prediction time.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{exp2_time_vs_reg}
	\caption{Results of experiment~2, plotted analogously
	to~\fref{fig:exp1-time-vs-reg}.}
	\label{fig:exp2-time-vs-reg}
\end{figure}

@@

In the third experiment we examine surrogate scaling properties by correlating
metrics of interest with training set size. Firstly, the results shown 
in~\fref{fig:scaling-r2} (and listed in~\tref{tbl:exp3-detailed-results-r2} in
the Appendix) suggest that the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based approaches
clearly outperform ANNs. This can be observed
particularly on sets of sizes up to~\num{6000}.

\begin{figure}[h]
	\centering
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{scaling_metric_r2}
		%\caption{Regression performance (as $R^2$)}
		%\label{fig:scaling-r2}
	\end{minipage}\hfill%
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{scaling_time_train}
		%\caption{Complexity (as~$\overline{t}_{\text{trn.}}$)}
		%\label{fig:scaling-trn}
	\end{minipage}\hfill%
	\begin{minipage}{0.333\textwidth}
		%\centering
		\includegraphics[width=\linewidth]{scaling_time_pred}
		%\caption{Complexity (as~$\overline{t}_{\text{pred.}}$)}
		%\label{fig:scaling-pred}
	\end{minipage}
	\caption{Various metrics collected during experiment 3 (scaling
	benchmark) displayed as a function of training set size.}
	\label{fig:scaling}
\end{figure}

Next, we examine scaling behaviour in terms of the mean training time (displayed
in~\fref{fig:scaling-trn} and listed in~\tref{tbl:exp3-detailed-results-t-train} in
the Appendix). Consistent with our expectation, the shortest times
were achieved by instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

Finally, we study scaling with respect to the mean prediction time (shown
in~\fref{fig:scaling-pred} and listed in~\tref{tbl:exp3-detailed-results-t-pred} in
the Appendix). Our initial observation is that all tested
families with the exception of previously mentioned instance-based
models offer desirable characteristics overall. Analogous to previous
experiments, GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following that, we notice a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

\subsubsection{Model Comparison}

@@

\begin{figure}[h]
	\centering
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model6}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model7}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model1}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model3}
	\end{minipage}

	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model4}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model5}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model2}
	\end{minipage}\hfill%
	\begin{minipage}{0.25\textwidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model8}
	\end{minipage}
	\caption{Regression performance of models 1-4 (row 1, from the left) and 5-8
		(row 2) trained in experiment~4 (model comparison), viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

@@

\begin{table}[h]
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\setlength\tabcolsep{2pt}
	{\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	}
	\caption{Results of experiment~4. Here, figures are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\tref{tbl:sampling-runs}).
		The best-performing metrics are highlighted in bold.}
	\label{tbl:exp4-detailed-results}
\end{table}

@@

\subsection{Results of Adaptive Sampling}
\label{sec:adaptiveres}
In order to test our QASS prototype, several functional toy theories for TBR were developed as alternatives to the expensive MC model. By far the most robust of these was the following sinusoidal theory with adjustable wavenumber parameter $n$:

\begin{equation}
	\text{TBR} = \frac{1}{|C|}\sum_{i \in C} \left[1 + \sin(2\pi n (x_i - 1/2)) \right]
\end{equation}

>>>>

\todo{Need any introduction here?}

\subsection{Decoupled Approach}
\label{sec:modelres}

@@

The results displayed in~\fref{fig:exp1-time-vs-reg} indicate that in the first,
simplified case GBTs clearly appear to be the most accurate as
well as the fastest surrogate family in terms of mean prediction time. Following
that, we note that ERTs, SVMs and ANNs also achieved satisfactory results with respect to both examined metrics.
While the remainder of tested surrogate families does not exhibit problems in
complexity, its regression performance falls below average.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{exp1_slice0}
	\includegraphics[width=0.9\linewidth]{exp1_slice1}
	\includegraphics[width=0.9\linewidth]{exp1_slice2}
	\caption{\label{fig:exp1-time-vs-reg}20~best-performing surrogates per each considered family, plotted in
		terms of complexity (as~$\overline{t}_{\text{pred.}}$) and regression
		performance (as~$R^2$) with 3~selected discrete feature assignments.}
\end{figure}

Comparing these results with those of the second, unrestricted experiment (shown
in~\fref{fig:exp2-time-vs-reg}), we observe that many surrogate families
consistently underperformed. The least
affected models appear to be GBTs, ANNs and ERTs, which are known to be capable of capturing relationships
involving mixed feature types that were deliberately withheld in the first
experiment. With only negligible differences, the first two of these families
appear to be tied for the best performance as well as the shortest prediction
time. We observe that ERTs and RBFs also
demonstrated satisfactory results, relatively outperforming the remaining surrogates in
terms of regression performance, and in some cases also in prediction time.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{exp2_time_vs_reg}
	\caption{\label{fig:exp2-time-vs-reg}Results of experiment~2, plotted analogously
	to~\fref{fig:exp1-time-vs-reg}.}
\end{figure}

@@

The results shown in~\fref{fig:scaling} suggest that in terms of regression
performance the most accurate families from the previous experiments
consistently maintain their relative advantage over others, even as we introduce
more training points. While such families achieve nearly comparable
performance on the largest dataset, in the opposite case tree-based approaches
clearly outperform ANNs. This can be observed
particularly on sets of sizes up to~\num{6000}.

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{scaling_metric_r2}
	\includegraphics[width=0.9\linewidth]{scaling_time_train}
	\includegraphics[width=0.9\linewidth]{scaling_time_pred}
	\caption{Results of the scaling benchmark displayed as a function of
	training set size. From top to bottom, regression performance (as $R^2$),
	mean training time and mean prediction time.}
	\label{fig:scaling}
\end{figure}

Consistent with our expectation, the shortest training times were achieved by
instance-based learning methods (e.g. KNN, IDW) that
are trained trivially at the expense of increased lookup complexity later during prediction.
Furthermore, we observe that the majority of tree-based algorithms also perform
and scale well, unlike RBFs and GPR which appear to behave superlinearly. We note that ANNs,
which are the only family to utilise parallelisation during training, show an
inverse scaling characteristic. Our conjecture is that this effect may be caused
by a constant multi-threading overhead that possibly dominates the training process
on relatively small training sets.

Finally, all tested families with the exception of previously mentioned instance-based
models offer desirable prediction times. Analogous to previous experiments,
GBTs, ABTs and ANNs appear to be tied, as they not only exhibit
comparable times but also similar scaling slopes. Following that, we note a
clear hierarchy of ERTs, SVMs, GPR and RBFs, trailed by IDW and KNNs.

\subsubsection{Model Comparison}
\todo{This section needs to be shortened.}

@@

\begin{figure}[h]
	\centering
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model6}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model7}
	\end{minipage}

	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model1}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model3}
	\end{minipage}

	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model4}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model5}
	\end{minipage}

	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model2}
	\end{minipage}\hfill%
	\begin{minipage}{0.5\linewidth}
		\centering
		\includegraphics[width=\linewidth]{exp4_model8}
	\end{minipage}

	\caption{Regression performance of models 1-4 (row 1, from the left) and 5-8
		(row 2) trained in experiment~4 (model comparison), viewed
		as true vs.~predicted TBR on a test set of a selected cross-validation
		fold. Points are coloured by density.}
	\label{fig:reg-performance}
\end{figure}

@@

\begin{table*}[t]
	\centering
	\sisetup{round-mode=places,round-precision=3,detect-weight=true,detect-family=true}
	\setlength\tabcolsep{2pt}
	{\scriptsize
		\begin{tabular}{lrrrrrrrr}
		\toprule
		{} & {} & \multicolumn{4}{c}{Regression performance} &
		\multicolumn{3}{c}{Complexity}\\
		\cmidrule(lr){3-6}
		\cmidrule(lr){7-9}
		Model & $|\mathcal{T}|$ & MAE [TBR] & $S$ [TBR] & $R^2$ [rel.] & $R^2_{\text{adj.}}$ [rel.]
						& $\overline{t}_{\text{trn.}}$ [\si{\milli\second}] &
		$\overline{t}_{\text{pred.}}$ [\si{\milli\second}] & $\omega$ [rel.]\\
		\midrule
		\input{tbl/table11}
		\bottomrule
		\end{tabular}
	}
	\caption{Results of experiment~4. Here, figures are reported over 5~cross-validation folds,
		$|\mathcal{T}|$~denotes cross-validation set size ($\times 10^3$)
		and $\omega$ is a relative speedup with respect to
		$\overline{t}_{\text{eval.}}=\SI{7.777049573054314}{\second}$
		measured in the MC TBR model during run~1 (see~\tref{tbl:sampling-runs}).
		The best-performing metrics are highlighted in bold.}
	\label{tbl:exp4-detailed-results}
\end{table*}

@@

\subsection{Adaptive Approach}\label{sec:adaptiveres}
\todo{This section needs to be shortened.}

In order to test our QASS prototype, several functional toy theories for TBR were developed as alternatives to the expensive MC model. By far the most robust of these was the following sinusoidal theory with adjustable wavenumber parameter $n$:

\begin{equation}
	\text{TBR} = |C|^{-1}\sum_{i \in C} \left[1 + \sin(2\pi n (x_i - 1/2)) \right]
\end{equation}
